All the concepts behind TrueSkill Through Time: the \proglang{Julia}, \proglang{Python} and \proglang{R} packages (Draft)

### 2020-12-15 article.tex
TTT: agrego correcciones de la última reunión

\en{Instead of selecting a single value as an estimate, we would like to quantify the ``relative certainty'' associated with each of the skill hypotheses.}
\es{En vez seleccionar un único valor como estimación, quisiéramos cuantificar la ``certidumbre relativa'' asociada a cada una de las hipótesis de habilidad.}
%
\en{The use of probability theory to handle this type of uncertainty (or beliefs), known as Bayesian inference, has been widely adopted not only because it has been shown to be a successful tool in practice~\citep{bishop2006-PRML}, but also because its rules have been proven to be the only ones that guarantee consistent reasoning~\citep{jaynes2003-bookProbabilityTheory,VanHorn2003}.}
\es{El uso de la teor\'ia de la probabilidad para tratar este tipo de incertidumbre (o creencias), conocida como inferencia Bayesiana, se ha adoptado ampliamente no sólo porque ha mostrado ser una herramienta exitosa en la práctica \citep{bishop2006-PRML}, sino también porque se ha demostrado que sus reglas son las únicas que garantizan un razonamiento consistente~\citep{jaynes2003-bookProbabilityTheory,VanHorn2003}.}

\en{All inference, no matter how complex, can be solved in terms of this two simple equations: the~\ref{eq:sum_rule} and the~\ref{eq:product_rule}.}
\es{Toda inferencia, no importa cuan compleja ella sea, puede ser resuelta a trav\'es de estas dos simples ecuaciones: la~\ref{eq:sum_rule} y la~\ref{eq:product_rule}.}

@@

\en{To reduce the uncertainty about the hidden skill using the information provided by the observed result and the described causal model, we need to solve:}

@@

 \underbrace{p(\overbrace{\text{\en{Skills}\es{Habilidades}}}^{\text{\en{Hidden}\es{Oculta}}}|\overbrace{\text{Result\es{ado}}}^{\text{Observ\en{ed}\es{ado}}}, \text{Model\es{o}})}_{\text{Posterior}} = \frac{\overbrace{P(\,\text{Result\es{ado}}\,|\,\text{\en{Skills}\es{Habilidades}}\,,\text{Model\es{o}})}^{\text{\en{Likelihood}\es{Verosimilitud}}}\overbrace{p(\text{\en{Skills}\es{Habilidades}})}^{\text{Prior}}}{\underbrace{P(\text{Result\es{ado}}\,|\,\text{Model\es{o}})}_{\text{Evidenc\en{e}\es{ia} o\en{r}\es{ predicci\'on a} prior \en{prediction}}}}

@@

\underbrace{p(s_i , \, r \, | \, s_j\, )}_{\hfrac{\text{\scriptsize Proportional}}{\text{\scriptsize posterior}}} = \underbrace{1-\Phi(0 \, |  \overbrace{s_i^{\textcolor{white}{2}} - \mu_j}^{\hfrac{\text{\en{Expected}\es{Diferencia}}}{\text{\en{difference}\es{esperada}}}} , \, \overbrace{2\beta^2 + \sigma_j^2}^{\hfrac{\text{\en{Total}\es{Incertidumbre}}}{\text{\en{uncertainty}\es{total}}}} )}_{\text{\en{Likelihood}\es{Verosimilitud}}} \,  \underbrace{\N(s_i \, | \, \mu_i,\, \sigma_i^2)}_{\text{Prior}} 
\en{Where the proportional posterior is just the unnormalized posterior, $p(s_i| r, s_j) = \frac{p(s_i, r|s_j)}{P(r)}$, and the likelihood is the cumulative density of all positive values of a Gaussian distribution centered at the difference between the hypothesis $s_i$ and the opponent's mean estimate $\mu_j$, with a variance that includes all uncertainties except the one of $s_i$.}
\es{El posterior no es m\'as que el posterior proporcional normalizado por la evidencia, $p(s_i| r, s_j) = \frac{p(s_i, r|s_j)}{P(r)}$, y la verosimilitud es la densidad acumulada de los valores positivos de una distribuci\'on Gaussiana centrada en la diferencia entre la hip\'otesis $s_i$ y la estimaci\'on media del oponente $\mu_j$, $\delta(s_i)$, con una varianza que incluye todas las incertidumbres salvo la de la propia hip\'otesis $s_i$, $\vartheta_i^2$. }

>>>>

\en{A superior solution, instead of selecting a single value as an estimate, involves quantifying the ``relative certainty'' associated with each of the skill hypotheses.}
\es{Una solución superior, a la de seleccionar una único valor como estimación, consiste en cuantificar la ``certidumbre relativa'' asociada a cada una de las hipótesis de habilidad.}
\en{This approach, known as Bayesian inference, which has proven successful in practice~\citep{bishop2006-PRML}, ensures consistent reasoning in handling this kind of uncertainty (plausible beliefs)~\citep{jaynes2003-bookProbabilityTheory,VanHorn2003}.}
\es{Este enfoque conocido como inferencia Bayesiana, que ha demostrado ser exitoso en la práctica~\citep{bishop2006-PRML}, garantiza un razonamiento consistente en el manejo de este tipo de incertidumbre (creencias plausibles)~\citep{jaynes2003-bookProbabilityTheory,VanHorn2003}.}
\en{All inference, no matter how complex it is, can be solved by the rules of probability: the~\ref{eq:sum_rule} and the~\ref{eq:product_rule}.}
\es{Toda inferencia, no importa cuan compleja sea, puede ser resuelta mediante las reglas de la probabilidad: la~\ref{eq:sum_rule} y la~\ref{eq:product_rule}.}

@@

\en{In our case, to reduce the uncertainty about the hidden skill using the information provided by the observed result and the described causal model, we need to solve:}

@@

 \underbrace{p(\overbrace{\text{\en{Skill}\es{Habilidad}$_i$}}^{\text{\en{Hidden}\es{Oculta}}}|\overbrace{\text{Result\es{ado}}}^{\text{Observ\en{ed}\es{ado}}}, \text{Model\es{o}})}_{\text{Posterior}} = \frac{\overbrace{P(\,\text{Result\es{ado}}\,|\,\text{\en{Skill}\es{Habilidad}$_i$}\,,\text{Model\es{o}})}^{\text{\en{Likelihood}\es{Verosimilitud}}}\overbrace{p(\text{\en{Skill}\es{Habilidad}$_i$})}^{\text{Prior}}}{\underbrace{P(\text{Result\es{ado}}\,|\,\text{Model\es{o}})}_{\text{Evidenc\en{e}\es{ia} o\en{r}\es{ predicci\'on a} prior \en{prediction}}}}

@@

\underbrace{p(\,s_i\, | \, r \, )}_{\text{Posterior}} \propto \underbrace{1-\Phi(0 \, |  \overbrace{s_i^{\textcolor{white}{2}} - \mu_j}^{\hfrac{\text{\en{Expected}\es{Diferencia}}}{\text{\en{difference}\es{esperada}}}} , \, \overbrace{2\beta^2 + \sigma_j^2}^{\hfrac{\text{\en{Total}\es{Incertidumbre}}}{\text{\en{uncertainty}\es{total}}}} )}_{\text{\en{Likelihood}\es{Verosimilitud}}} \,  \underbrace{\N(s_i \, | \, \mu_i,\, \sigma_i^2)}_{\text{Prior}} 
\en{Where the normalized posterior can be found dividing the right hand by the evidence ${P(r)}$.}
\es{Donde el el posterior normalizado se obtiene dividiendo el lado derecho por la evidencia, $P(r)$.}
\en{The likelihood is just the cumulative density of all positive values of a Gaussian distribution centered at the difference between the hypothesis $s_i$ and the opponent's mean estimate $\mu_j$, with a variance that includes all uncertainties except the one of $s_i$.}
\es{La verosimilitud es la densidad acumulada de los valores positivos de una distribuci\'on Gaussiana centrada en la diferencia entre la hip\'otesis $s_i$ y la estimaci\'on media del oponente $\mu_j$, $\delta(s_i)$, con una varianza que incluye todas las incertidumbres salvo la de la propia hip\'otesis $s_i$, $\vartheta_i^2$. }
